---
title: "C lacinia Modeling"
author: "Maggie Vincent"
date: "2024-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Preference measurements

We measured 

```{r}

# load in the data from our subset file -- this subset just means I've split apart the oviposition data and cleaned some things up

# I also added family to this data file and added a few variations of times -- rounded times as well as times in seconds

choiceDF <- as.data.frame(read.csv("~/Work/chlosyne_lacinia/mating_trials_2023/data_code/2023_choice_trials_subset_HH_removed.csv"))

```



**First, I will model how parental treatment impacts the time before copulation in seconds.**

```{r}
m1 <- lm(time_before_seconds ~ as.factor(combination_treatment), data = choiceDF)

summary(m1)

res1 <- resid(m1)

hist(res1)

# I am going to try log-transforming time_before_seconds to see if the residuals are better; I'd like the median to be closer to 0 with a more even distribution overall

m2 <- lm(log(time_before_seconds) ~ as.factor(combination_treatment), data = choiceDF)

summary(m2)

res2 <- resid(m2)

hist(res2)

# they're not perfect but the median is much better so I'll stick with the log-transformed variable for this model

```

# log(time before copulation) ~ combination treatment

F2,11 = 3.114; p-value = 0.0847.



**Now I will model the effect of mating pair treatment on the duration of copulation in seconds.**

```{r}

m3 <- lm(duration_seconds ~ as.factor(combination_treatment), data = choiceDF)

summary(m3)

# I am happy with these residuals, we get a median of 0 with min/max of -468/676

# I am not going to transform anything here

res3 <- resid(m3)

hist(res3)

```


# duration of copulation ~ combination treatment

F2,11 = 0.4678, p = 0.6383




##### Fitness Proxies


Measurements: number of clutches, number of hatched offspring, number of adult offspring, and average pupation time in days.


```{r}
# load in the fitness df

fitnessDF <- as.data.frame(read.csv("~/Work/chlosyne_lacinia/mating_trials_2023/data_code/choice_f2_fitness_data_HH_removed.csv"))
```



**Modeling the effect of parental treatment on clutch number**


```{r}

m4 <- lm(clutch_number ~ as.factor(cross_type), data = fitnessDF)

summary(m4)

res4 <- resid(m4)

hist(res4)

# these residuals look normal, I am not going to transform anything

```


# clutch number ~ parental treatments

F2,10 = 0.004931, p = 0.9951



**Modeling the effect of parental treatment on the number of hatching offspring**


```{r}

m5 <- lm(hatch_number ~ as.factor(cross_type), data = fitnessDF)

summary(m5)

res5 <- resid(m5)

hist(res5)

# residuals are not centered at 0 (median = -8), but they are evenly distributed

```


# hatch number ~ parental treatment

F2,9 = 0.4464, p = 0.6533



**Modeling the effect of parental treatment on the number of adult offspring**


```{r}
m6 <- lm(adult_number ~ as.factor(cross_type), data = fitnessDF)

summary(m6)

res6 <- resid(m6)

hist(res6)

# the residuals are not very normal, I am going to transform adult number and try again

fitnessDF_noNA <- na.omit(fitnessDF)

m7 <- lm(log(adult_number) ~ as.factor(cross_type), data = fitnessDF_noNA)

summary(m7)

res7 <- resid(m7)

hist(res7)

# not amazing, but better, we will use this one
```


# log(adult number of offspring) ~ parental treatment

F2,7 = 0.8406, p = 0.4708



**Modeling the effect of parental treatment on the average pupation time of offspring**


```{r}
m8 <- lm(average_pupation_time ~ as.factor(cross_type), data = fitnessDF)

summary(m8)

res8 <- resid(m8)

hist(res8)

# residuals look good
```


# average pupation time of offspring ~ parental treatment

F2,8 = 0.8438, p = 0.4651







