---
title: "C Lacinia Mapping"
author: "Maggie Vincent"
date: "2024-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Mapping practice

I am following the documentation from https://eriqande.github.io/rep-res-web/lectures/making-maps-with-R.html

```{r}
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
```

The packages maps and mapdata have outlines and ggmap and ggplot2 are so I can plot maps.

```{r}
usa <- map_data("usa")
ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group = group)) + 
  coord_fixed(1.3)
```

This is a basic, filled in map of the USA. I can make aesthetic adjustments.

```{r}
ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = NA, color = "red") + 
  coord_fixed(1.3)
```

Another example of aesthetic adjustments, this time with a fill color.

```{r}
gg1 <- ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = "violet", color = "blue") + 
  coord_fixed(1.3)
gg1
```


### Adding points to the map.

```{r}
labs <- data.frame(
  long = c(-122.064873, -122.306417), #give the long of the two points
  lat = c(36.951968, 47.644855), #give the lat of the two points
  names = c("SWFSC-FED", "NWFSC"), #here they are naming the two points
  stringsAsFactors = FALSE
  )

gg1 + 
  geom_point(data = labs, aes(x = long, y = lat), color = "black", size = 5) +
  geom_point(data = labs, aes(x = long, y = lat), color = "yellow", size = 4)

#the first geom_point line is to make the whole dot and the second line is to fill it in with yellow
```

### Now just mapping one state, the example given is California

```{r}

ca_df <- subset(states, region == "california")

ca_county <- subset(counties, region == "california")

ca_base <- ggplot(data = ca_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")

ca_base + 
  geom_polygon(data = ca_county, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA)  # get the state border back on top

```


Okay I want to subset to just a map of Texas.

```{r}
states <- map_data("state")

tex_map <- subset(states, region == "texas")

counties <- map_data("county")
tx_county <- subset(counties, region == "texas")

# map of Texas, no county lines
tx_base <- ggplot(data = tex_map, mapping = aes(x = long, y = lat)) +
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "white")


#adding county lines
tx_base + geom_polygon(data = tx_county, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA)  # get the state border back on top


```

### Adding coordinates

```{r}

colony_1_coordinates <- data.frame(long = c(-99.19156
, -99.817, -99.817, -99.589, -99.472, -99.472, -99.472, -98.663), lat = c(28.66565, 28.493, 28.493, 28.124, 27.877, 27.876, 27.877, 26.316), names = c("pop 1", "pop 2", "pop 3", "pop 4", "pop 5", "pop 6", "pop 7", "pop 8"),
  stringsAsFactors = FALSE)



# now I will add colony 1 coordinates to my base map of Texas and save that as an object

tx_col1_only <- tx_base + 
  geom_point(data = colony_1_coordinates, aes(x = long, y = lat), color = "black", size = 2)

# add colony 2 coordinates to tx_col1_only but pick a different color so it's clear

colony_2_coordinates <- as.data.frame(read.csv("~/Work/chlosyne_lacinia/mating_trials_2023/data_code/colony_2_coordinates.csv"))
                         

c2_coords <- colony_2_coordinates %>%
  rename(long = longitude,
         lat = latitude)

c_lac_collection_map <- tx_col1_only + 
  geom_point(data = c2_coords, aes(x = long, y = lat), color = "orange", size = 2)


```



