---
title: "C. lacinia Mate Choice"
author: "Maggie Vincent"
date: "2024-02-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# I have mate/no mate data from the choice and no choice trials. I want to see if they differ from each other more than expected by chance. 

```{r}
mate_nomate_all <- as.data.frame(read.csv("~/Work/chlosyne_lacinia/mating_trials_2023/data_code/all_mate_no_mate_CL.csv"))

choice_chi_data <-  as.data.frame(read.csv("~/Work/chlosyne_lacinia/mating_trials_2023/data_code/chi_treatment_choice_numeric.csv"))
```


# I have a dataframe with the mate/no mate counts for the choice and no choice trials.

chisq.test(x=mate_nomate_all$choice, p=mate_nomate_all$no_choice, rescale.p = TRUE)

# The mate/no mate counts are significantly different between the choice and no choice trials.


``` {r}
# I want to see if the choice is different than it would be if females picked randomly


chisq.test(choice_chi_data$female_high, choice_chi_data$female_low, correct=FALSE)

# In the two chi squared tests above, the observed data does not differ from what we would predict if random mating was occurring. 
```

Kyle had some concerns about not doing a proper pairwise comparison in the above test, so I did some reading. I found a package called "rcompanion" with the function "pairwiseNominalIndependence" so I am going to try that below.



```{r}

library(rcompanion)

pairwiseNominalIndependence(choice_chi_data,
                            fisher = FALSE,
                            gtest = FALSE,
                            chisq = TRUE,
                            stats = TRUE
                            )




```









